---
title: "sinadef"
output: html_document
date: "2024-04-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Nos conectamos al dataset:
Columns:

Education: The educational qualifications of employees, including degree, institution, and field of study.

Joining Year: The year each employee joined the company, indicating their length of service.

City: The location or city where each employee is based or works.

Payment Tier: Categorization of employees into different salary tiers.

Age: The age of each employee, providing demographic insights.

Gender: Gender identity of employees, promoting diversity analysis.

Ever Benched: Indicates if an employee has ever been temporarily without assigned work.

Experience in Current Domain: The number of years of experience employees have in their current field.

Leave or Not: a target column

```{r}
url <- "https://raw.githubusercontent.com/chelocoder/ProyectoR/main/Employee.csv"
datos <- read.csv(url)
head(datos)
```
```{r count-missing-data, results='markup'}
# Conteo de datos faltantes en cada columna
missing_data_counts <- sapply(datos, function(x) sum(is.na(x)))

# Mostrando el conteo de datos faltantes
missing_data_counts

```


```{r}
summary(datos)
```


```{r}
# Identificando variables numéricas
numeric_vars <- sapply(datos, function(x) is.numeric(x) || is.integer(x))
numeric_variables <- names(datos)[numeric_vars]

# Identificando variables categóricas
categorical_vars <- sapply(datos, function(x) is.factor(x) || is.character(x) || is.logical(x))
categorical_variables <- names(datos)[categorical_vars]

# Mostrando las variables numéricas y categóricas
list(numeric = numeric_variables, categorical = categorical_variables)
```
```{r barplot-joining-year-ggplot, fig.cap="Gráfico de Barras de JoiningYear con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un data frame con los conteos de cada año
joining_year_data <- as.data.frame(table(datos$JoiningYear))

# Crear el gráfico de barras usando ggplot2
ggplot(joining_year_data, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  ggtitle("Gráfico de Barras de JoiningYear") +
  xlab("Año de Ingreso") +
  ylab("Frecuencia")
```
```{r barplot-paymenttier-ggplot, fig.cap="Gráfico de Barras de PaymentTier con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un data frame con los conteos de cada categoría de PaymentTier
payment_tier_data <- as.data.frame(table(datos$PaymentTier))

# Crear el gráfico de barras usando ggplot2
ggplot(payment_tier_data, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "coral") +
  ggtitle("Gráfico de Barras de PaymentTier") +
  xlab("Tier de Pago") +
  ylab("Frecuencia")

```


```{r histogram-age-ggplot, fig.cap="Histograma de Age con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un histograma usando ggplot2
ggplot(datos, aes(x = Age)) +
  geom_histogram(binwidth = 1, fill="blue", color="black") +
  ggtitle("Histograma de Age") +
  xlab("Edad") +
  ylab("Frecuencia")

```
```{r boxplot-age-ggplot, fig.cap="Box Plot de Age con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Creando un box plot usando ggplot2
ggplot(datos, aes(y = Age)) +
  geom_boxplot(fill = "lightgreen") +
  ggtitle("Box Plot de Age") +
  ylab("Edad")

```


```{r barplot-experience-ggplot, fig.cap="Gráfico de Barras de ExperienceInCurrentDomain con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un data frame con los conteos de cada valor de ExperienceInCurrentDomain
experience_data <- as.data.frame(table(datos$ExperienceInCurrentDomain))

# Crear el gráfico de barras usando ggplot2
ggplot(experience_data, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  ggtitle("Gráfico de Barras de ExperienceInCurrentDomain") +
  xlab("Años de Experiencia en el Dominio Actual") +
  ylab("Frecuencia")

```
```{r barplot-leaveornot-ggplot, fig.cap="Gráfico de Barras de LeaveOrNot con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un data frame con los conteos de cada categoría de LeaveOrNot
leave_data <- as.data.frame(table(datos$LeaveOrNot))

# Crear el gráfico de barras usando ggplot2
ggplot(leave_data, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "salmon") +
  ggtitle("Gráfico de Barras de LeaveOrNot") +
  xlab("Decisión") +
  ylab("Frecuencia")

```


```{r}
# Identifying character variables
char_vars <- sapply(datos, class) == "character"

# Extracting unique values for each character variable
unique_values <- lapply(datos[, char_vars, drop = FALSE], unique)

# Displaying the unique values
unique_values

```
```{r barplot-education-ggplot, fig.cap="Gráfico de Barras de Education con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un data frame con los conteos de cada categoría de Education
education_data <- as.data.frame(table(datos$Education))

# Crear el gráfico de barras usando ggplot2
ggplot(education_data, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "turquoise") +
  ggtitle("Gráfico de Barras de Education") +
  xlab("Nivel de Educación") +
  ylab("Frecuencia")

```
```{r barplot-city-ggplot, fig.cap="Gráfico de Barras de City con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un data frame con los conteos de cada ciudad
city_data <- as.data.frame(table(datos$City))

# Crear el gráfico de barras usando ggplot2
ggplot(city_data, aes(x = reorder(Var1, -Freq), y = Freq)) +
  geom_bar(stat = "identity", fill = "cadetblue") +
  ggtitle("Gráfico de Barras de City") +
  xlab("Ciudad") +
  ylab("Frecuencia") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotating x-axis labels for better visibility

```

```{r barplot-gender-ggplot, fig.cap="Gráfico de Barras de Gender con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un data frame con los conteos de cada género
gender_data <- as.data.frame(table(datos$Gender))

# Crear el gráfico de barras usando ggplot2
ggplot(gender_data, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "orchid") +
  ggtitle("Gráfico de Barras de Gender") +
  xlab("Género") +
  ylab("Frecuencia")

```
```{r barplot-everbenched-ggplot, fig.cap="Gráfico de Barras de EverBenched con ggplot2"}
# Cargar la librería ggplot2
library(ggplot2)

# Crear un data frame con los conteos de cada categoría de EverBenched
everbenched_data <- as.data.frame(table(datos$EverBenched))

# Crear el gráfico de barras usando ggplot2
ggplot(everbenched_data, aes(x = Var1, y = Freq)) +
  geom_bar(stat = "identity", fill = "lightcoral") +
  ggtitle("Gráfico de Barras de EverBenched") +
  xlab("Ha Sido Bencheado") +
  ylab("Frecuencia")

```
```{r setup, include=FALSE}
# Install GGally if it's not already installed
if (!require("GGally")) install.packages("GGally", dependencies = TRUE)

# Load the necessary libraries
library(ggplot2)
library(GGally)


```

```{r convert-factors}
# Converting 'LeaveOrNot' to a factor if it's not already
datos$LeaveOrNot <- as.factor(datos$LeaveOrNot)
# Optionally, convert other categorical variables
datos$PaymentTier <- as.factor(datos$PaymentTier)

# Now, subset the dataset with the required variables
selected_variables <- datos[, c("Age", "JoiningYear", "ExperienceInCurrentDomain", "PaymentTier", "LeaveOrNot")]
```
```{r setup-libraries, include=FALSE}
# Install corrplot if it's not already installed
if (!require("corrplot")) install.packages("corrplot", dependencies = TRUE)

# Load the library
library(corrplot)

```

```{r pairplot-leaveornot-fixed, fig.cap="Pairplot with LeaveOrNot as Hue"}
# Creating the pairplot using GGally
ggpairs(selected_variables, aes(color = LeaveOrNot), 
        lower = list(continuous = wrap("points", alpha = 0.5, size = 1)),
        diag = list(continuous = wrap("barDiag")),
        upper = list(continuous = wrap("cor", size = 3)),
        title = "Pairplot with LeaveOrNot as Hue")

```


```{r compute-correlation, results='markup'}
# Subsetting the dataset to include only numerical variables
numerical_data <- datos[sapply(datos, is.numeric)]

# Calculating the correlation matrix
cor_matrix <- cor(numerical_data, use="pairwise.complete.obs")  # Handling missing values by considering complete cases

# Display the correlation matrix
cor_matrix

```
```{r visualize-correlation, fig.cap="Correlation Matrix Visualization"}
# Visualizing the correlation matrix
corrplot(cor_matrix, method="circle", type="upper", order="hclust",
         tl.col="black", tl.srt=45)  # Rotate labels for better readability
```

